The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Adjusted to Eslevier style. Work on text to make it fit in 10 pages
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1745 4525493e-7705-40b1-a816-d608a930855b

\section{Data Aggregation System\label{DAS}}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. We carefully analyzed user
requests and queries placed into the DBS, their patterns, frequencies and latency. 
As a result we created a data discovery service \cite{DD}. Originally it used MVC
architecture and later, being re-factored as pure presentation layer on top of 
the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
we took away relations and details on underlying schema from end-users. 
This was achieved by usaging Dijkstras shortest path algorithm 
to establish necessary table joins based on provided information in user queries and
knowledge of foreign key relationships from auto-loaded database schema.
The DBS-QL keys exposed to end-users were choosen based on a jargon used by
physicists. They're internally mapped into database entitites and attributes, 
e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Due to this set of requirements, we rule out the choice of RDMS as DAS back-ends
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistency in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexies, including inner objects and embedded arrays,
and auto-sharding. Our prelimiary benchmarks shown that it can sustain
the desired load and size for our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of the
following components:
core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregated results;
request service to handle user requests;
mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map,
data parser(s) suitable for data-service output format(s) into DAS common
data format (JSON) and was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}\footnote{We used regular expression 
patterns, evaluated during run time, to validate provided set of input 
paremters and DAS keys.}:
\begin{verbatim}
{"api": {"params": {"name": ""}, "name": "CMSNametoSE"}, 
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", 
              "das_key": "site"}], 
 "_id": "4ae1f8dfe2194e4581000014", 
 "system": "sitedb", 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", 
              "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
we dealt with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS caching system}
The DAS caching system was used to store data-service results and aggregated
information in form of DAS records. As it shown in figure \ref{DAS_cache} 
the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
and aggregate data over there. The independent set of daemons, DAS robots, were
constantly monitor Analytics DB for most popular requests and pre-fetch most
common data into DAS cache.

Each user request placed to the system was parsed and set of selection
and condition keys were identified. They were mapped into set of
data service APIs and their input parameters. DAS
cache was checked for a queries with a similar set of selection and condition
keys used earlier.\footnote{Since DAS didn't hold
any data by default we cannot rely on data being present in a cache.
Moreover if we look-up data first, we cannot tell if it was complete
set or not, since earlier queries may retrieve only portion of the requested data.}
If the superset was found a data were
retrieved from the cache, otherwise a data-service APIs were
invoked and the results were collected and aggregated inside of DAS cache.

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys known to DAS to describe common
entities of our data. For example, {\it dataset, file, block, run, event}. Those
names were used by our users in their daily work operations. So, by typing
{\it file} we didn't perform a string matching to ``file'' string, but rather
look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
For intance, for {\it dataset, block, file} the default attribute 
was a {\it name}, while for {\it run} and {\it event} the default attribute was 
a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
{\it run=123456}, which means run number 123456. At the same time a queyr 
key, e.g. {\it site} meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as {\it run=123456} instruct system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

@@

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS data files. In former case, DBS 
stored information about files and their relationship to other 
physics objects, such as run and luminosity, while PhEDEx kept
information about the file location, e.g. details about sites 
holding copy of the files. Upon a user request to find a file, 
we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects. For instance, since both data-services returns in their
output a file name we used as a common key during aggregation.
This merged results represented a set of DAS records. Each DAS
record contains a standard DAS header, which identify look-up time,
request url, input parameters and API method as well as response 
version, expiration time and checksum of the output. Usage of this 
information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc. Here is an example of DAS record representing
aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
and {\it CMStoSiteName}, with API results merged under the {\it site} key.
\begin{verbatim}
{"_id": "4ae1e7ffe2194e4243000003", 
 "site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
 "das": [{"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 0.95439004898071289, "expire": 1256362175.7692671, 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSAMName", 
          "timestamp": 1256318975.7692671, "system": ["sitedb"], 
          "api": ["CMStoSAMName"], "version": "", "selection_keys": []}, 
         {"qhash": ["ed2b73067794d443ed56756d9b94f7bc"], 
          "ctime": 1.2158100605010986, 
          "api": ["CMStoSiteName"], 
          "url": "https://cmsweb.cern.ch/sitedb/json/index/CMStoSiteName", 
          "timestamp": 1256318974.7774379, "system": ["sitedb"], 
          "expire": 1256362174.7774379, "version": "", "selection_keys": []}
        ]
}
\end{verbatim}

@@

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system and data-service;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
Figure \ref{db_size} shows the current size of meta-data from aforementioned 
data-services.
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{db_size.pdf}
\caption{
Current size of meta-data for CMS data-services participated in DAS, before
official data-taking. 
%We anticipate to collect $\sim$500GB of meta-data
%per year during normal LHC operations.
}
\label{db_size}
\end{figure}

At the moment we don't have exact numbers for amount of data which will stored 
in those data-service and therefor accessible in DAS. Although we project to
accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
operations.\footnote{This number is based on tests 
performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{llllll}\hline
\hline

System & Format & Size & Records & Elapsed time & Elapsed time \\
& & & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
system we show total number of fetched records 175884 together with
number of records merged in DAS 170131. The total number of DAS records 
were calculated as number of aggregated records plus left over, non-matched records,
from both systems.}
\label{DAS_benchmark}
\end{table*}

Please note that initial DAS implementation was done in Python \cite{Python}
and all tests were performed on x64 Linux platform with 16GB of RAM where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
of DAS server was 67\% in average and RAM usage was below 35MB for
all tests. All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert notations used by data-service to DAS ones for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
on their common key (block name in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.

As you can see from Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

@@

\section*{References}
\begin{thebibliography}{9}
\bibitem{DBXplorer}
S. Agrawal, S. Chaudhuri, G. Das,
``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
ICDE 2002, pp. 5-16.

@@

\bibitem{CMSDataModel} 
C. Grandi, D.Stickland, L.Taylor et al.,
``The CMS Computing Model'',
CERN-LHCC-2004-035/G-083 (2004);
A. Fanfani et. al., 
``Distributed Analysis in CMS'', 
to be published in Journal of Grid Computing.

@@

\bibitem{DBS07} 
A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'',
Journal of Physics: Conference Series 119, 072013, (2008).

@@

\bibitem{CRAFT09}
https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo

\bibitem{Python} Python programming language, http://www.python.org

>>>>

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose design, implementation and
security policies are left up to their developers. Therefore DAS does not require
a data preservation and transaction capabilities. Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
ability to place queries, similar to SQL, without specific knowledge of 
underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Based on this set we rule out the choice of RDMS as DAS back-end(s)
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistency in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexies, including inner objects and embedded arrays,
and auto-sharding. Our prelimiary benchmarks shown that it can sustain
the desired load and size of our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregated results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
Analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}:
%\footnote{We used regular expression 
%patterns, evaluated during run time, to validate provided set of input 
%paremters and DAS keys.}:
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
to deal with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions ($conds$) 
and selection keys ($skeys$)
\item Check cache for query hash, if found: look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define method to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previos step
\begin{verbatim}
if cache_has(row.match_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
We did use a bulk insert operations to increase throughput. The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. for 
\verb+file+ record it is a \verb+file.name+.
\item Get results from the cache
\end{enumerate}
The data was stored in a form of DAS records into the cache, while
queries were recorded into Analytics DB for further analysis. We
also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

%The DAS caching system was used to store data-service results and aggregated
%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
%and aggregate data over there. The independent set of daemons, DAS robots, were
%constantly monitor Analytics DB for most popular requests and pre-fetch most
%common data into DAS cache.

%Each user request placed to the system was parsed and set of selection
%and condition keys were identified. They were mapped into set of
%data service APIs and their input parameters. DAS
%cache was checked for a queries with a similar set of selection and condition
%keys used earlier.\footnote{Since DAS didn't hold
%any data by default we cannot rely on data being present in a cache.
%Moreover if we look-up data first, we cannot tell if it was complete
%set or not, since earlier queries may retrieve only portion of the requested data.}
%If the superset was found a data were
%retrieved from the cache, otherwise a data-service APIs were
%invoked and the results were collected and aggregated inside of DAS cache.

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys known to DAS to describe common
entities of our data. For example, {\it dataset, file, block, run, event}. Those
names were used by our users in their daily work operations. So, by typing
{\it file} we didn't perform a string matching to ``file'' string, but rather
look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
For intance, for {\it dataset, block, file} the default attribute 
was a {\it name}, while for {\it run} and {\it event} the default attribute was 
a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
{\it run=123456}, which means run number 123456. At the same time a query 
key, e.g. {\it site} meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as {\it run=123456} instruct system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

@@

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS data files. In former case, DBS 
stored information about files and their relationship to other 
physics objects, such as run and luminosity, while PhEDEx kept
information about the file location, e.g. details about sites 
holding copy of the files. Upon a user request to find a file, 
we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects. For instance, since both data-services returns in their
output a file name we used as a common key during aggregation.
This merged results represented a set of DAS records. Each DAS
record contains a standard DAS header, which identify look-up time,
request url, input parameters and API method as well as response 
version, expiration time and checksum of the output. Usage of this 
information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc. Here is an example of DAS record representing
aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
and {\it CMStoSiteName}, with API results merged under the {\it site} key.
\begin{verbatim}
{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
 "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
          "timestamp": 1256318975, "system": ["sitedb"]},
         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
          "timestamp": 1256318974, "system": ["sitedb"]]}
\end{verbatim}

@@

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
%Table \ref{db_size} shows the current size of meta-data from aforementioned 
%data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

%\begin{table*}[hbt]
%\centering
%\begin{tabular}{llllll}\hline
%\hline

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment we don't have exact numbers for amount of data which will stored 
in those data-service and therefor accessible in DAS. Although we project to
accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
operations.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{llllll}\hline
\hline
System & Format & Size & Records & Elapsed time & Elapsed time \\
& & & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
system we show total number of fetched records 175884 together with
number of records merged in DAS 170131. The total number of DAS records 
were calculated as number of aggregated records plus left over, non-matched records,
from both systems.}
\label{DAS_benchmark}
\end{table*}

Please note that initial DAS implementation was done in Python 
and all tests were performed on x64 Linux platform with 16GB of RAM where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
of DAS server was 67\% in average and RAM usage was below 35MB for
all tests. All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert notations used by data-service to DAS ones for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
on their common key (block name in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
As you can see from Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

@@

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with BibTeX database:

%\bibliographystyle{elsarticle-num}
%\bibliography{<your-bib-database>}

%% Authors are advised to use a BibTeX database file for their reference list.
%% The provided style file elsarticle-num.bst formats references in the required Procedia style

%% For references without a BibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\section*{References}
\begin{thebibliography}{00}
\bibitem{DBXplorer}
S. Agrawal, S. Chaudhuri, G. Das,
``DBXplorer: A System for Keyword-Based Search over Relational Databases'',
ICDE 2002, pp. 5-16.

@@

\bibitem{CMSDataModel} 
C. Grandi, D.Stickland, L.Taylor et al.,
``The CMS Computing Model'',
CERN-LHCC-2004-035/G-083 (2004).
%A. Fanfani et. al., 
%``Distributed Analysis in CMS'', 
%to be published in Journal of Grid Computing.

\bibitem{iRODS}
The Integrated Rule-Oriented Data System,
http://www.irods.org/

\bibitem{OpenArchive}
Open Archives Initiative,
http://www.openarchives.org/

@@

\bibitem{DBS07} 
A. Dolgert, V. Kuznetsov, C. Jones, D. Riley, 
``A multi-dimensional view on information retrieval of CMS data'',
J. Phys. Conf. Ser, 119, 072013 (2008).

@@

%\bibitem{CRAFT09}
%https://twiki.cern.ch/twiki/bin/view/CMS/CRAFT09AnalysisInfo

%\bibitem{Python} Python programming language, http://www.python.org

%\bibitem{DBSearch}
%D. Konopnicki, O. Shmueli,
%``Database-Inspired Search'', 
%Proc. of the 31st VLDB Conference, 2005.
\end{thebibliography}
\end{document}
